---
title: "2020 Final Practice 5pm"
output:
  pdf_document: default
  keep_tex: yes
  html_document:
    df_print: paged
subtitle: Economemtrics 
fontsize: 12pt
linestretch: 1.5
---

**5a.**
```{r}
library(AER)
data("CigarettesSW")

head(CigarettesSW)

CigarettesSW$rprice <- CigarettesSW$price / CigarettesSW$cpi
CigarettesSW$rtax <- CigarettesSW$tax / CigarettesSW$cpi
CigarettesSW$rtaxs <- CigarettesSW$taxs / CigarettesSW$cpi
CigarettesSW$rincome <- (
    CigarettesSW$income / (CigarettesSW$cpi * CigarettesSW$pop)
)
```

**5b.**
```{r}
CigarettesSW$eighty_five <- ifelse(
    CigarettesSW$year == 1985, 1, 0
)

model <- lm(
    log(packs) ~ 1 + log(rprice) + log(rincome) + eighty_five,
    data = CigarettesSW
)

test_homo <- coeftest(model, vcov = vcovHC(model, type = "HC0"))
print(test_homo)
```

**5c.**

No

**5d.**

```{r}
test_hetero <- coeftest(model, vcov = vcovHC(model, type = "HC1"))
print(test_hetero)
```

**5e.**

People with lower incomes may be correlated with wore socioeconomic circumstances which may increase frequencey of smoking.

**5f.**

```{r}
model_2sls <- ivreg(
    log(packs) ~ 1 + log(rprice) + log(rincome) + eighty_five |
        log(income) + rtax + rtaxs + eighty_five,
    data = CigarettesSW
)

summary(model_2sls, diagnostics = TRUE)
```

**5g.**

No evidence that IVs are endogenous

**5h.**

Null hypothesis of exogeneity of errors is significant at 10%. Combination with simultaneity problem, we should use 2SLS.

**5i.**

Demand is less elastic with respect to price. Demand is less lastic with respect to income.

**5j.**

```{r}
coeftest(model_2sls, vcov = vcovHC(model, type = "HC1"))
```